### 2012-06-15 humanoid.tex
Add a picture to the small space controllability proof. Minor corrections

\begin{tikzpicture}[y=0.55pt, x=0.55pt,yscale=-1, inner sep=0pt, outer sep=0pt]
  \path[cm={{0.91292191,0.40813428,-0.40813428,0.91292191,(116.6153,-98.710192)}},draw=black,fill=black,miter
  limit=4.00,fill opacity=0.621,line width=2.400pt]
  (373.7564,122.0117)arc(0.000:180.000:58.841385 and
  23.739)arc(-180.000:0.000:58.841385 and 23.739) -- cycle;
\path[cm={{0.91292191,0.40813428,-0.40813428,0.91292191,(-6.875857,3.3152152)}},draw=black,fill=black,miter
  limit=4.00,fill opacity=0.621,line width=2.400pt]
  (373.7564,122.0117)arc(0.000:180.000:58.841385 and
  23.739)arc(-180.000:0.000:58.841385 and 23.739) -- cycle;
\path[shift={(26.616754,-41.712121)},draw=black,fill=black,miter limit=4.00,fill
  opacity=0.621,line width=2.400pt] (275.7716,94.7376)arc(0.000:180.000:8.838835
  and 8.586)arc(-180.000:0.000:8.838835 and 8.586) -- cycle;
\path[draw=black,dash pattern=on 1.60pt off 0.80pt,line join=miter,line
  cap=butt,miter limit=4.00,line width=0.800pt] (187.8884,281.1107) --
(402.9370,93.8920) node [right=5] {$y$-axis};
\path[draw=black,line join=miter,line cap=butt,line width=0.870pt]
  (284.9572,56.9063) -- (202.5850,114.7252) -- (254.4490,172.5441) --
  (232.6622,243.4361);
\path[draw=black,line join=miter,line cap=butt,line width=0.722pt]
  (302.9315,55.9207) -- (371.2797,74.1552) -- (344.6998,94.3611) --
  (353.6560,136.7942);
\path[fill=black] (245,42.362179) node[above right] (text3828) {CoM};
\path[draw=black,dash pattern=on 1.60pt off 0.80pt,line join=miter,line
  cap=butt,miter limit=4.00,line width=0.800pt] (292.5714,63.0765) --
  (292.5714,219.2525);
\path[fill=black] (270,136.6479) node[above right] (text3842) {$z_c$};
\path[draw=black,line join=miter,line cap=butt,line width=0.800pt, style=<->]
(299.6429,281.2908) -- (361.7857,227.3622); 
\node(L) at (335,260) [right] {$L$};
\path[draw=black,line join=miter,line cap=butt,line width=0.800pt,style=<->]
  (339.2857,155.9336) -- (372.5000,127.0050);
\node (l0) at (365,145) {$l$};

\end{tikzpicture}